\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-19 paper.tex
Rephrase, re-organize discussion, taking into account Rachid's contributions

Also, we would like to mention that, after some research (see appendix B
of~\cite{Lemaignan2012a} for a detailed discussion)  we have decided to
represent neither the planning domain nor the resulting plans in the knowledge
base: the planning domain (with task pre- and postconditions) is stored in a
specific format, outside of the central declarative knowledge repository, and
the plans are directly communicated to the robot controller. This leads to
independent declarative and procedural knowledge stores.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Robot Execution Control}
\label{sect|ctrl}

@@

One of the main role of {\sc Shary} is to control the elaboration and execution
of shared plans~\cite{Grosz1996,Clark1996,Kemp2007}. This means essentially
contex-based refinement and robot actions execution as well as monitoring of
those achieved by its human partner~\cite{Breazeal2003}. The challenge is to
build such abilities in a generic way, and to provide several levels of
parametrization allowing to adapt to various environments, and various levels of
involvement of the robot ranging from teammate behavior to assistant or
proactive helper.  Depending on this, the robot controller invokes the various
human-aware planners and will react to events detected at ORO level.

@@

This last section tries to clarify the ``takeaway'' message: what are the challenges
that human-robot interaction brings to artificial intelligence, and what is the
current state-of-the-art in this field?

We first restate the challenge of \emph{embodied cognition} in general and list
what we see as the main decisional issues that need to be tackled to push
forward human-robot interaction. We then reflect on the importance of knowledge
handling in robotic architectures that deal with complex semantics and put into
perspective our choice of the RDF formalism.  Finally, we briefly re-discuss two
facets of HRI: the importance of \emph{modelling the humans}, and the difficulty
of \emph{acting in presence of humans}.

@@

This leads to a tight coupling between the symbolic and the geometric realms:
while AI at its origins was mostly a matter of symbolic models, it has been
since recognised that not only the mind is not a purely abstract system,
disconnected from the physical world, but even more, cognition fundamentally
relies on its relation to the physical world (so-called \emph{embodied
cognition}). Varela~\cite{Varela1992} is one of the main discoverers of these
mechanisms, and coined the concept of \emph{enactivism} as the theoretical
framework that study the links between cognition, embodiment and actions.

@@

As we have seen, grounding is implemented at different levels in our
architecture. The main source of grounded symbolic knowledge is the situation
assessment module, {\sc Spark}.  Not only it builds symbolic facts from spatial
reasoning, but it also relies on (limited) temporal reasoning to track the world
state and build explanations to interpret unexpected perceptions (like a
suddenly disappearing object).

Because {\sc Spark} is also able to track humans, which enables in turn
perspective-aware spatial reasoning, it produces grounded symbolic knowledge
for interacting agents as well. This is a typical \emph{embodied} cognitive
skill.

Grounding also occurs during verbal and non-verbal communication. The {\sc
Dialogs} module grounds new concepts introduced by the speaker by asking questions
until it can attach the new concept to concepts already present in the
knowledge pool (typically, if I ask the robot ``bring me a platypus'', the
robot asks ``what is a platypus''. I may answer ``a mammal'' and the robot will
ask ``what is a mammal?'' and so on until it anchors the new concepts to ones
it already knows). Non verbal interaction (like gestures) is also partially
grounded: we have presented how pointing, for instance, is used by the robot to
ground \emph{``this''} or \emph{``that''} to the pointed artifact.

Because only objects marked with tags are currently recognised (in our typical
lab experiments, it may be a dozen of them), the robot grounding mechanisms work
in a small-sized closed world. This assumption obviously greatly facilitates
the task. It remains to be seen if systematic marking of manipulated objects
(for instance with RFID tags) is a realistic path.\footnote{It may be the case if the
robot is expected to be used in a well-defined, slowly changing environment.
Real-time recognition of large libraries of objects is another path which may
be viable soon~\cite{Dean2013Fast}.}

We see however that we have only scratched the surface of a generic grounding
capability. Context, cultural background, ``naive physics'' knowledge, emotional
state of the human are some of the numerous other determinants that need to be
accounted for when grounding human-robot interaction.

@@

How can we restate the main challenges, in particular in term of decision
making, that human-robot interaction brings to artificial intelligence?

\fxfatal{TDB}

@@

@@

\subsection{Bringing the Humans into the Equation}

\subsubsection{Modelling the Humans}

\fxfatal{Re-organise, rephrase, complement}

In our implementation, perspective taking is tightly connected to the symbolic
knowledge models, and since our knowledge base allows for storage of one
knowledge model per agent, we have been able to endow the robot with a simple
theory of mind (as explained in section~\ref{sect|tom}): we explicitly model
what the robot knows about its partners in a symbolic way. This knowledge is
then re-used in different places, to correctly interpret what the human says,
or to plan tasks that are actually doable for the human.

The cognitive model that the robot builds for the agent it interacts with is
still simple and mostly focused on geometric features (\emph{Who sees what? What are
our relative positions?} etc.). Extending this knowledge with more subtle
perceptions (emotional state for instance) remains to be done.

We should also mention that modelling of human behaviours also appear in our
architecture at the level of task and motion planning with a set of
\emph{social rules}, including proxemics, action legibility or time
optimization.

\subsubsection{Acting in Presence of Humans}

HRI offers new challenges to planning by adding a supplementary constraint: the
synthesis of plans that are not only feasible but legible (in terms of their
intention) and compliant with human preferences. This constraint is often
ill-defined, dynamic and context-dependent.  HATP, by co-planning for the robot and
the human, and modelling social rules as planning parameters, is a first step in
this direction.

\fxfatal{What are the next steps?}

\paragraph{Joint actions with humans}\label{sec:soa}
\fxfatal{too many citation -> synthesis needed!}
The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasise that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realise it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they
all have shown benefits of this approach. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels.

>>>>

Also, we would like to mention that, after some research (see appendix B
of~\cite{Lemaignan2012a} for a detailed discussion)  we have decided to
represent neither the planning domain nor the resulting plans in the knowledge
base: the planning domain (with task pre- and postconditions) is stored in a
specific format, outside of the central declarative knowledge repository, and
the plans are directly communicated to the robot controller. This leads to
independent declarative and procedural knowledge stores.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Robot Execution Control}
\label{sect|ctrl}

@@

One of the main role of {\sc Shary} is to control the elaboration and execution
of shared plans~\cite{Grosz1996,Clark1996,Kemp2007}. This means essentially
contex-based refinement and robot actions execution as well as monitoring of
those achieved by its human partner~\cite{Breazeal2003}. The challenge is to
build such abilities in a generic way, and to provide several levels of
parametrization allowing to adapt to various environments, and various levels of
involvement of the robot ranging from teammate behaviour to assistant or
proactive helper.  Depending on this, the robot controller invokes the various
human-aware planners and will react to events detected at ORO level.

@@

This last section tries to clarify the ``takeaway'' message: what are the
challenges that human-robot interaction brings to artificial intelligence?  We
first restate the challenge of \emph{embodied cognition} in general and
summarise what we see as the main decisional issues that need to be tackled to
push forward human-robot interaction. Finally, we reflect on the importance of
knowledge handling in robotic architectures that deal with complex semantics and
put into perspective our choice of the RDF formalism.

@@

This leads to a tight coupling between the symbolic and the geometric realms:
while AI at its origins was mostly a matter of symbolic models, it has been
since recognised that not only the mind is not a purely abstract system,
disconnected from the physical world, but even more, cognition fundamentally
relies on its relation to the physical world (so-called \emph{embodied
cognition}). Varela~\cite{Varela1992} is one of the main discoverers of these
mechanisms, and coined the concept of \emph{enactivism} as the theoretical
framework that study the links between cognition, embodiment and actions.  This
has since been thoroughly studied in robotics and artificial intelligence
(Pfeifer and Bongard~\cite{pfeifer2007body} is one of the reference).

@@

As we have seen, grounding is implemented at different levels in our
architecture. The main source of grounded symbolic knowledge is the situation
assessment module, {\sc Spark}. It builds symbolic facts from spatial reasoning,
and also relies on (limited) temporal reasoning to track the world state and
build explanations to interpret unexpected perceptions (like an object that
suddendly disappears). Because {\sc Spark} also tracks humans, which enables
perspective-aware spatial reasoning, it can produce as well grounded symbolic
knowledge for the agents it interacts with. This is a typical \emph{embodied}
cognitive skill.

Grounding also occurs during verbal and non-verbal communication. The {\sc
Dialogs} module grounds new concepts introduced by the speaker by asking
questions until it can attach the new concept to concepts already present in the
knowledge pool (typically, if I ask the robot ``bring me a platypus'', the robot
asks ``what is a platypus''. I may answer ``a mammal'' and the robot will ask
``what is a mammal?'' and so on until it anchors the new concepts to ones it
already knows). Embodied interactions (like gestures) are also taken into
account at this level: we have presented how pointing, for instance, is used by
the robot to ground \emph{``this''} or \emph{``that''} to the pointed artifact.

Note that because only objects marked with 2D fiducial markers are currently
recognised (in our typical lab experiments, it may be a dozen of them), the
robot's grounding mechanisms only work in a small-sized closed world. This
assumption greatly facilitates the task, and we see that we have only scratched
the surface of a generic grounding capability. Context, cultural background,
``naive physics'' knowledge, emotional state of the human are some of the
numerous other determinants that need to be accounted for when grounding
human-robot interaction.

@@

With this perspective on \emph{embodied cognition} in mind, it seems appropriate
to restate the key challenges brought by human-robot interaction to artificial
intelligence.

\subsubsection{The W-questions}

Our context, where a robot has to achieve a task together with
humans, raises issues to be tackled by the robots' decisional components, and
that we can summarises as ``W-questions'': \emph{What}, \emph{Who}, \emph{When},
\emph{Where} and \emph{How}?

\emph{What to do next}, at different levels of abstractions and while taking
into account not only the current state but also the long term goals, is the
basic question for an intelligent robot. Its is made more complex here by having
to deal with the partially observable physical and mental state of the human
partner, and by the extended set of possible actions.

\emph{Who should act now} is of key importance and also needs to be
decided upon by the robot. It is sometimes expected as an intelligent behaviour
for the robot to wait and let its human partner act instead. Correct management
of \emph{turn-taking} leads to various decisional challenges.

\emph{When to perform a given action} is enriched here by the human-context
since it has to take into account the human, his/her needs, his/her rhythms and
pace, and his/her mental state. While performing its share of the task, the
robot has to produce signals directed towards the human and to respond to
signals produced back at the proper pace.

\emph{Where to perform an action} plays an important role as well: the choice is
not trivial and might need elaborated decision. The robot is expected to take
into account effort sharing, visibility of its action by the human, disturbance
or discomfort induced by its action.

\emph{How to perform an action}, finally, needs to be reflected upon by the
robot: several options to perform an action or to achieve a goal are often
available, and selecting one is a non-trivial decision problem. Cost-based
planners are one possible approach: they search for plans that satisfy an
acceptable cost in terms of \emph{acceptability} or \emph{legibility}, for
instance.

These five questions should not be considered independently from each others and
often require, on the contrary, to be dealt with in a single decision step. The
human-aware task and motion planners which we have built, are instances of
systems which have been designed to deal with such intricate decision issues.

\subsubsection{Putting the Humans into Equations}

The correlate of these five \emph{W-questions} is the issue the \emph{human
models}: taking appropriate decisions with and in presence of humans require
appropriate models of the human. Models that tell the robot what the human
\emph{can do}, \emph{would like to do}, \emph{knows}, \emph{could infer}, etc.

While the task of describing all the (dynamic!) human models that are useful to
robots is immense (if doable at all), we claim that it is possible to devise and
use such models in limited, but still interesting and useful, contexts such as
collaborative human-robot objects manipulation, fetch-and-carry and associated
activities in a domestic or professional environment.

In our implementation, perspective taking, for instance, is tightly connected to
the symbolic knowledge models, and since our knowledge base allows for storage
of one knowledge model per agent, we have been able to endow the robot with a
simple theory of mind (as explained in section~\ref{sect|tom}): we explicitly
model what the robot knows about its partners in a symbolic way. This knowledge
is then re-used in different places, to correctly interpret what the human says,
or to plan tasks that are actually doable for the human.

The cognitive model that the robot builds for the agents it interacts with
remains today simple and mostly focused on geometric features and affordances
(\emph{who sees what? what are our relative positions? what is reachable to
whom?}  etc.). Extending this knowledge with more subtle perceptions (emotional
state for instance) remains to be done.

Motion and action execution also requires human models, and the one we use
embeds human preferences and physical constraints that need to be accounted for
when synthesizing robot motion or elaborating robot plans. This includes
proxemics (human-robot distance) and associated issues (visibility) but also
legibility and acceptability criterias expressed in terms of \emph{social rules}
that the elaborated plans should satisfy.

%
%\paragraph{Joint actions with humans}\label{sec:soa}
%\fxfatal{too many citation -> synthesis needed!}
%
%The human presence brings new requirements for robot's abilities both
%at the functional and at the deliberative levels~\cite{Klein2004}. The
%topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
%navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
%in presence of humans as well as perception of human
%activities~\cite{Breazeal2001,Burger2008}. Also, when
%interacting with humans, robots need to incorporate communication and
%collaboration abilities. Several theories dealing with
%collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasise that
%collaborative tasks have specific requirements compared to individual
%ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realise it, they must show their
%commitment to the goal during execution, etc. Several robotic systems
%have already been built based on these
%theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they
%all have shown benefits of this approach. They have also shown how
%difficult it is to manage turn-taking between communication partners
%and to interleave task realization and communication in a generic
%way. Finally, today only few
%systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
%account at all levels.

@@

As thorougly presented in this article, we have built the decisional
capabilities of our robots around this idea of explicit knowledge manipulation.
We finally briefely comment on this design choice, our implementation based on RDF,
and the limitations that we perceive.

@@
